
---
title: "Very useful tools in R/Python for Geospatial Analysis"
---
  
  In this document, I will explain some very useful and easy 
  tools that I use in R for geospatial analysis and for spatial object
  manipulation (e.g shapefiles, raster files,
  spatial databases, etc.). I also going to present some libraries and 
  commands that I usually use in my-day-to-day code work for cleaning 
  and dataset manipulation (Stata alike)
  
  For all this I'm going to use mainly examples of public databases
  from the Peruvian context (the country that I come from). This is going to be 
  a very short post, but I'm going to try to include great variety of tools
  at least for every basic analysis (e.g open spatial object, modify them, plot
  maps, polygons, points, etc.)
  

  ```{r include = FALSE}
    # set working directory - this does not appear in the document
    rm(list = ls())
    
    setwd("~/Dropbox/paulo_RA/Narcos/databases") # Paulo's MacbookPro

    library(rgdal)
  ```
  Open shapefiles:
  
  ```{r}
    mun <- readOGR("./municipios/update", "BAS_LIM_DISTRITOS")
    # An easly inspect some objects:
    plot(mun)
    
  ```
