
---
title: "Very useful tools in R/Python for Geospatial Analysis"
output: github_document
---
  
  In this document, I will explain some very useful and easy 
  tools that I use in R for geospatial analysis and for spatial object
  manipulation (e.g shapefiles, raster files,
  spatial databases, etc.). I also going to present some libraries and 
  commands that I usually use in my-day-to-day code work for cleaning 
  and dataset manipulation (Stata alike)
  
  For all this I'm going to use mainly examples of public databases
  from the Peruvian context (the country that I come from). This is going to be 
  a very short post, but I'm going to try to include great variety of tools
  at least for every basic analysis (e.g open spatial object, modify them, plot
  maps, polygons, points, etc.)
  

  ```{r include = FALSE}
  
    rm(list = ls()) # to clear your working spaces

    # These are the libraries that we are going to use for this small post
    
    library(sp)
    library(rgdal)
    library(readstata13)
    library(raster)
    
    # For graph: 
    library(tidyverse)
    library(broom) 
    
  ```
  
## The basics
  
  Open shapefiles:
  
  ```{r}
    mun <- readOGR(paste0("~/Dropbox/paulo_RA/Narcos/databases/municipios/update"), 
                          "BAS_LIM_DISTRITOS") # rgdal library
                          
    # An easly inspect some objects:
    plot(mun)
    
  ```
  
  Of course, we are able not only to open a shp but even a latlong file and 
  transform it to a GIS kind of object. See for example:
  
  ```{r}
    enaho <- read.dta13(paste0("~/Dropbox/BDatos/ENAHO/2020/737-Modulo01/",
      "enaho01-2020-100.dta")) # Kind of large dataset around 50k observations
    
    # However we can easly handle this as an spatial object:
    
    coordinates(enaho) <- ~ longitud + latitud
  
    # A very important thing is to identify coords systm. For usual lat/long 
    # the following works:
    
    proj4string(enaho) <- CRS("+init=epsg:4326") # Here we are defining the coord syst.
      # sp library
    
    plot(enaho) # And we can plot it 
  ```
  
  Now we have in our R work enviroment two spatial objects, a polygon object, 
  Peruvian municipalities, and a point object, Peruvian household for the Annual
  National Survey (ENAHO). Of course we can print this together in a graph, but 
  first we need to check if both spatial objects have the same coords system
  
  ```{r}
    print(crs(enaho))
    print(crs(mun)) ## crs from raster library
  ```
  
  As you may see, the municipality dataset is in UTM coord system (projected in 
  meters), and the household survey is in standard latlong coordinates. We can
  easily transform everything in UTM coords (it is going to be useful later on)
  as follow:
  
  ``` {r}
    enaho <- spTransform(enaho, crs(mun))
    
    # And of course we can plot it together in a graph 
    plot(mun) # And we can plot it 
    plot(enaho, add = TRUE, col='red', pch=1, cex = .2) # And we can plot it
    
  ```
  
  