
---
title: "Very useful tools in R for Geospatial Analysis"
output: github_document
---
  
  In this document, I will explain some very useful and easy 
  tools that I use in R for geospatial analysis and for spatial object
  manipulation (e.g shapefiles, raster files,
  spatial databases, etc.). I also going to present some libraries and 
  commands that I usually use in my-day-to-day code work for cleaning 
  and dataset manipulation (Stata alike)
  
  For all this I'm going to use mainly examples of public databases
  from the Peruvian context (the country that I come from). This is going to be 
  a very short post, but I'm going to try to include great variety of tools
  at least for every basic analysis (e.g open spatial object, modify them, plot
  maps, polygons, points, etc.)
  

  ```{r include = FALSE}
  
    rm(list = ls()) # to clear your working spaces

    # These are the libraries that we are going to use for this small post
    
    library(sp)
    library(rgdal)
    library(readstata13)
    library(raster)
    
    # For graph: 
    library(tidyverse)
    library(broom) 
    
  ```
  
## Level 0: Setup, open and inspect spatial object
  
  Open shapefiles:
  
  ```{r}
    mun <- readOGR(paste0("~/Dropbox/paulo_RA/Narcos/databases/municipios/update"), 
                          "BAS_LIM_DISTRITOS") # rgdal library
                          
    # An easly inspect some objects:
    plot(mun)
    
  ```
  
  Of course, we are able not only to open a shp but even a latlong file and 
  transform it to a GIS kind of object. See for example:
  
  ```{r}
    enaho <- read.dta13(paste0("~/Dropbox/BDatos/ENAHO/2020/737-Modulo01/",
      "enaho01-2020-100.dta")) # Kind of large dataset around 50k observations
    
    # However we can easly handle this as an spatial object:
    
    coordinates(enaho) <- ~ longitud + latitud
  
    # A very important thing is to identify coords systm. For usual lat/long 
    # the following works:
    
    proj4string(enaho) <- CRS("+init=epsg:4326") # Here we are defining the coord syst.
      # sp library
    
    plot(enaho) # And we can plot it 
  ```
  
  Now we have in our R work enviroment two spatial objects, a polygon object, 
  Peruvian municipalities, and a point object, Peruvian household for the Annual
  National Survey (ENAHO). Of course we can print this together in a graph, but 
  first we need to check if both spatial objects have the same coords system
  
  ```{r}
    print(crs(enaho))
    print(crs(mun)) ## crs from raster library
  ```
  
  As you may see, the municipality dataset is in UTM coord system (projected in 
  meters), and the household survey is in standard latlong coordinates. We can
  easily transform everything in UTM coords (it is going to be useful later on)
  as follow:
  
  ``` {r}
    enaho <- spTransform(enaho, crs(mun))
    
    # And of course we can plot it together in a graph, and add one more 
    # shapeline of lines 
    
    # Rivers in Peru:
    
    rivers <- readOGR(paste0("~/Dropbox/BDatos/Shape/PER_wat"), 
                          "PER_water_areas_dcw") 
                          
    rivers <- spTransform(rivers, crs(mun)) # just in case is in other crs 
    
    
    plot(mun, lwd = .1) # And we can plot it 
    plot(rivers, add = TRUE, border = "blue") # add lines
    plot(enaho, add = TRUE, col='red', pch=1, cex = .1) # add points 
    
  ```
  
  Of course these are not the most rad graphs, but they accomplish their
  main purpose inspect the element. I will do another post with more
  stylish figures ;)
    
## Level 1: Basic operations
  
  We've already discussed the initial steps for any geographic analysis. Now I
  am going to present some of the most basic operations that we can perform on
  spatial objects. The good thing is that actually, we can do everything 
  that we used to do on a regular datasets (subsetting, math operations, 
  bysorts, conditionals, etc...)

  ``` {r}
  
    # Subsample
    mun <- mun[mun$NOMBPROV == 'LIMA',] # easily subsampling
    
    # Include example of dummy gen here and plot // --> to do
    
    # Manipulating variables
    enaho$prov_id = substr(enaho$ubigeo,1,4) # creating an prov ID
    enaho$household_id = paste0(enaho$ubigeo, enaho$conglome, enaho$vivienda, 
                          enaho$hogar) # creating a household ID
    
    # We of course can quickly check what we did (in 10 random rows):
    
    sample_n(enaho@data[,c("ubigeo", "conglome", "vivienda", 
                           "hogar", "prov_id", "household_id")], 10) 
                           
    # Notice the use of @data in the previous command, sometimes, we need to 
    # be more specific with R (but in most of the cases
    # we can treat spatial dataframe just as a regular dataframe
    
    # Our household ID is unique of course 
    print(length(enaho$household_id)) # Number of rows 
    print(length(unique(enaho$household_id))) # Number of unique rows 
    
    # We can check the number of municipalities and provinces too 
    
    print(length(unique(enaho$ubigeo)))
    print(length(unique(enaho$prov_id)))
    
    # We are going to subsample our household database too, in order to match with 
    # the mun database 
    
    enaho <- enaho[enaho$prov_id == '1501',] # LIMA province
    
    # And we can check our subsampling:
    
    plot(mun, lwd = .5) # And we can plot it 
    plot(enaho, add = TRUE, col='red', pch=1, cex = .1) # add points
    
  ```
  
  And of course, as we can transform regular dataframes with lat/long to 
  spatial objects, we are able to turnback them to regular dataframes (this
  can be useful in several cases, e.g when you would like to reduce estimation
  time). Moreover, we can perform other more 'complicated' data operations, like 
  merge with other data information.
  
  
## Level 2: Spatial operations - spatial merge, aggregations, grid generation, and more


## Level 3: Elevation and raster objects


  